Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-03-05 WSSSPE2_report/WSSSPE2_review.tex
working in section 7 - in progress

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\subsection{Papers}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\begin{itemize}
\item Mario {Rosado de Souza} and Robert Haines and Caroline Jay. Defining
Sustainability through Developers' Eyes: Recommendations from an Interview
Study~\cite{wssspe2_rosada_de_souza}

@@

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\subsection{Papers}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\begin{itemize}
\item Jordan Adams, Sai Nudurupati, Nicole Gasparini, Daniel Hobley, Eric
Hutton, Gregory Tucker and Erkan Istanbulluoglu. Landlab: Sustainable Software
Development in Practice ~\cite{wssspe2_adams}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Credit \& Incentives} \label{sec:credit}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo{Dan (OR VOLUNTEER) to lead writing of this section}

@@

This group, with just three papers but good interest and participation from attendees,
focused on the institutional, social, and cultural mechanisms encourage the creation
and maintenance of shared software, in the context of what now exists but focusing
on what mechanisms are desired and how we might achieve them.

@@

The group had some initial discussion ...

The group then broke into three subgroups
\begin{itemize}
\item Citation ecosystems
\item Taxonomy of contributorship/Guidelines for software for tenure review
\item Prizes
\end{itemize}

citation ecosystems sub-breakout: \url{http://tinyurl.com/l4tc6a4}
%https://docs.google.com/document/d/1SOja7k0mXwvOB2tQD3ZW3O38Qv4NwfA21zdMQ9JlUQc/edit

Taxonomy of Contributorship/Guidelines for Software for Tenure Review
sub-breakout: \url{http://tinyurl.com/mx58tmx}
%https://docs.google.com/document/d/1sDZLaBR2elZ9Tj6eLIwEtf5VqQ_x_AN_SetKtJlFUpI/edit

Prizes sub-breakout: \url{http://tinyurl.com/mjok4o5}
%https://docs.google.com/document/d/1foSppeHQeQ-qOjd-Rp_CUCNy8JMLxd_vrCDMmqvG2-w/edit 
(but as of 2/6, nothing is here and need to look at main breakout doc - Suresh
is moving text, I think)

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\subsection{Papers}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\begin{itemize}
\item James Howison. Retract bit-rotten publications: Aligning incentives for
sustaining scientific software~\cite{wssspe2_howison}

@@

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\subsection{Papers}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\begin{itemize}
\item Jakob Blomer, Dario Berzano, Predrag Buncic, Ioannis Charalampidis,
Gerardo Ganis, George Lestaris and Ren\'{e} Meusel. The Need for a Versioned
Data Analysis Software Environment~\cite{wssspe2_blomer}

@@

\begin{itemize}
\item Thomas Clune, Michael Rilee and Damian Rouson. Testing as an Essential
Process for Developing and Maintaining Scientific Software~\cite{wssspe2_clune}

>>>>

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\subsection{Papers}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
The papers that were discussed in the Exploring Sustainability group are:
\begin{itemize}
\item Mario {Rosado de Souza} and Robert Haines and Caroline Jay. Defining
Sustainability through Developers' Eyes: Recommendations from an Interview
Study~\cite{wssspe2_rosada_de_souza}

@@

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\subsection{Papers}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
The papers that were discussed in the Software Development Experiences group are:
\begin{itemize}
\item Jordan Adams, Sai Nudurupati, Nicole Gasparini, Daniel Hobley, Eric
Hutton, Gregory Tucker and Erkan Istanbulluoglu. Landlab: Sustainable Software
Development in Practice ~\cite{wssspe2_adams}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Credit \& Incentives} \label{sec:credit}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo{Dan leading writing of this section}

@@

This group, with just three papers but a large amount of interest and participation from attendees,
focused on the institutional, social, and cultural mechanisms encourage the creation
and maintenance of shared software, in the context of what now exists but focusing
on what mechanisms are desired and how we might achieve them.

@@

In the first discussion session, this group decided to break into two smaller groups,
each independently working through the same general topic: credit and incentives.

\subsubsection{First breakout discussion: Group A}
The first sub-group discussed around the current system for credit and current incentives,
which it called ``hacking the incentive structure''.  The group considered four potential
points of leverage:

First, that we currently have systems that collect information, and these could be modified
to collect different information, then map that information to actions. We could
initially build a proof-of-concept for a new use of a given system, then determine
what actions would be needed to make this use more common.

Second that we could create entirely new systems, perhaps because the existing
systems are too tied to what they measure, and modifying them is not practical.

Third that we could change academic culture, rather than worrying about the systems.
This was mostly focused on citations, because they matter for hiring, promotion,
and tenure decisions.  The group discussed how we could weight the citations
within papers better than we now do? How could we identify the five citations
that really matter for a paper, distinguishing them from the related works and
general background that are also cited.  Perhaps we could break these out in
the reference list, working with publishers to implement this.  Or maybe we
could also break out categories of citations, such as the most important software used,
the previous publication that we are building from, the data that we actually used, etc.
The Moore Foundation's award in data science was given as an example, asking
proposers: What are the five canonical citations that are most important to your work?
This would be a way of giving credit and assigning importance to these works,
differently from how we just count citations today.  A possible action that the group
discussed was conducting a longitudinal study of most useful \{software, data, etc.\}
in a discipline

Fourth, that we could change the ways funders make decisions, and use these funding
policies as incentives.

After this discussion, group A brainstormed about incentives,
with the following items suggested:
\begin{itemize}
\item running programming contests, creating bounties for contributing to open source software, etc.
\item augment author lists to give credit to people who don’t now get credit (and making them machine readable)
\item developing a microcitation standard and mechanism (for both software and data)
\item developing a well-defined standard for author contribution -- what level of contribution rises to the level of authorship?
\item leveraging social media for citation and reviewing of content -- then using social media to bring more people into the review process than is traditional
\item determining where else software can be cited and recorded (e.g., acknowledgments sections of papers)
\item developing a taxonomy of contributors, tied to places that these metrics are already stored (e.g., ORCID)
\item making metadata more easy to add for software, creating incentive for providing software metadata -- note that this cannot be centralized
\item creating something like the h-index that tenure committees can make use of - simplify a way of measuring and documenting the overall credit given to an individual over different projects
\item thinking about publishing software vs journal - software does not have to be novel
\item determining guidelines for recommending software characteristics for tenure - perhaps draft guidelines then get ACM or IEEE agreement
\end{itemize}

\subsubsection{First breakout discussion: Group B}
\note{writing in progress}
Discussion Extracts

Who to incentivize: 
Those in science (to do better, more shareable, more sustainable work).
Those in industry (but are interested in science).

Giving credit in the traditional way:  software papers (JORS?).
Difficulty of adding new authors to publications
 - original authors disproportionately benefit
 - how can you add authors?
Can we control the end-users?
 - Yes, but we prefer something not so "confrontational"
Versioning.  How can we add that to publications? To add on people.  JSON-LD percentages.
John Towns: What do you want to encourage?
 - "good computational work" to advance their careers.
 - People to realize that I'm doing good software.
 - Empirical Q: Most of the developers are not going into traditional scientific careers.
    - Why are people going to Industry?  Money? Or working without worrying about careers paths
Portable reputation.
Trying to support career paths of "those in the grey area (between science and software - in development of software)"
 - How do you ensure that those "engineers" do not block the path "back into science".  Force the PIs to make programmers co-authors.

Prizes: reframe software as "instrument building"
	will prizes be good or bad, how can we
Converting reputation by joining apache software foundation.
	Google etc.  Inviting ASF and open source people to contribute to scientific code (at least on questions of overlapping

Template for assessing scientific contributions made through software.

making decisions about who to fund from funder perspective - who did what in the project?

someone writes paper in some area, adds references at end to make sure it’s clear that the paper fits in a field, but maybe don't think about how we use citations to give credit - how to rank/say what citations really matter, different from those that just explain the field (required background)

\subsubsection{Group merger and redivision}

After the first breakout session, the groups A and B came together and discussed 
a rollup of the ideas from the subgroups at a high level:
\begin{itemize}
\item citation ecosystem - traces of usage (metrics)
\item taxonomy of contributorship, understanding roles
\item prizes
\item new metrics (for people's activities in software)
\item guidelines for evaluating scientific contribution through software (perhaps using new metrics)
\end{itemize}

In the remaining discussion sessions, the group chose to split into three subgroups to discuss
a version of these topics: citation ecosystems,
taxonomy of contributorship/guidelines for software for tenure review, and prizes.

\subsubsection{Remaining breakout discussions: Citation ecosystems}
\note{writing to be done}

\url{http://tinyurl.com/l4tc6a4}
%https://docs.google.com/document/d/1SOja7k0mXwvOB2tQD3ZW3O38Qv4NwfA21zdMQ9JlUQc/edit

\subsubsection{Remaining breakout discussions: Taxonomy of contributorship/guidelines for software for tenure review}
\note{writing to be done}

sub-breakout: \url{http://tinyurl.com/mx58tmx}
%https://docs.google.com/document/d/1sDZLaBR2elZ9Tj6eLIwEtf5VqQ_x_AN_SetKtJlFUpI/edit

\subsubsection{Remaining breakout discussions: Prizes}
\note{writing to be done}

\url{http://tinyurl.com/mjok4o5}
%https://docs.google.com/document/d/1foSppeHQeQ-qOjd-Rp_CUCNy8JMLxd_vrCDMmqvG2-w/edit

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\subsection{Papers}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
The papers that were discussed in the Citation \& Incentives group are:
\begin{itemize}
\item James Howison. Retract bit-rotten publications: Aligning incentives for
sustaining scientific software~\cite{wssspe2_howison}

@@

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\subsection{Papers}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
The papers that were discussed in the Reproducibility \& Reuse \& Sharing group are:
\begin{itemize}
\item Jakob Blomer, Dario Berzano, Predrag Buncic, Ioannis Charalampidis,
Gerardo Ganis, George Lestaris and Ren\'{e} Meusel. The Need for a Versioned
Data Analysis Software Environment~\cite{wssspe2_blomer}

@@

The papers that were discussed in the Code Testing \& Code Review group are:
\begin{itemize}
\item Thomas Clune, Michael Rilee and Damian Rouson. Testing as an Essential
Process for Developing and Maintaining Scientific Software~\cite{wssspe2_clune}
